---
title: "Assignment06"
author: "Kaylie Evans"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction
The goal of this RMD is to take one of the New York Times APIs, construct an interface in R to read in the JSON data and transform it into an R DataFrame. The NYT API that was chosen is the Books API and the data to be taken out of that is the current best sellers list for graphic books and manga.

## Import Libraries
```{r library-import}
library(tidyverse) 
library(jsonlite)
library(httr)
```

## Import Web API
```{r api-import}
# this API key was assigned to my account on Times Developers
api_key <- "q3Ei9rbguhYDwWjqyiAC1sw19y5wSgx9"

# url from API site
books_url <- paste0("https://api.nytimes.com/svc/books/v3/lists/current/graphic-books-and-manga.json?api-key=", api_key)

# response with httr's GET
response <- GET(books_url)

# translate json encoding
json_data <- content(response, "text", encoding = "UTF-8")
parsed_data <- fromJSON(json_data)

#take a look at the data to see what transformation needs to be done
glimpse(parsed_data)
```

## Transform JSON Format to R Data Frame
From the above glimpse(), it is clear that there are 2 lists that need to be unnested: 

.. ..$ isbns               :List of 15
.. ..$ buy_links           :List of 15

The Results list contains another list called Books. The data below comes directly from the Books list so the Results list does not need to be unnested.
```{r transformation-flatten}
# take the subgroup books from results as that has the data we are interested in
books_df <- parsed_data$results$books

# flatten the nested structures within the 'isbns' column
books_df <- unnest(books_df, isbns)

# flatten the nested structures within the 'buy_links' column
books_df <- unnest(books_df, buy_links)

#check the data frame
glimpse(books_df)

```


## Conclusion

The data frame is translated correctly in R, containing 144 rows and 28 columns. More cleaning and transformation would need to be done for a simple analysis; however, that is out of the scope of this markdown.
